Lifting CDCL to Numerical Domains

### 2016-09-27 domains.tex
octagon domain description

The commonly used library for numerical abstract domains  
is the APRON C library~\cite{apron}.  This library is 
used for the static analysis of the numerical variables 
of a program by abstract interpretation. APRON provides a 
C API interface to various abstract domains and libraries 
such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
{\em Linear Equalities} library.  The aim of such analysis is 
to compute invariants over numerical variables in the 
program~\cite{se2011}.

@@

The join $(\join_{O})$ of two octagons, $\langle (v_i \times v_j \rightarrow N_1),
(v_i \times v_j \rightarrow N_2) \rangle$, is not necessarily an octagon 
and is computed by taking piece-wise maximum of bounds of corresponding 
octagonal inequalities, $\langle (v_i \times v_j \rightarrow max(N_1, N_2) \rangle$.
However, the meet $(\meet_{O})$ of two octagons, $\langle (v_i \times v_j \rightarrow N_1),
(v_i \times v_j \rightarrow N_2) \rangle$, is always an octagon and is 
computed by taking piece-wise minimum of bounds of corresponding 
octagonal inequalities, $\langle (v_i \times v_j \rightarrow min(N_1, N_2)$.

For building program analyzers using octagon domains, the domain also 
supports few operators like {\em widening ($\nabla$)} and {\em closure ($*$)}.
The closure operator is often used to reduce the degree of over-approximation 
resulting from the join operation. Closure always produces unique octagonal 
inequalities, but it is very expensive to compute because of cubic complexity 
in the number of program variables.   
The octagon domain is a relational numberic abstract domain that  
Domain:
Operators:
Semantics of elements:

\subsection{Conditions for Absract Domains}
One of the requirements for a clause learning SAT solver 
is that each element of the partial assignments domain 
needs to have a decomposition into precisely complementable 
elements.  This property of the underlying domain helps to 
guide the search away from the conflicting region of the 
search space.

Thus, clause learning SAT solvers requires a weak complementation property
of the abstraction.  However, numerical abstract domain,  such as 
intervals and octagons are intersections of complementable half-spaces. 
APRON library is specialized for the abstract interpretation 
framework which does not require complementation 
operator.  To this end, we implement our own template-based 
octagon domain which supports complementation operator.

>>>>

@@

The join $(\join_{O})$ of two octagons, $\langle (v_i \times v_j \rightarrow N_1),
(v_i \times v_j \rightarrow N_2) \rangle$, is not necessarily an octagon 
and is computed by taking piece-wise maximum of bounds of corresponding 
octagonal inequalities, $\langle (v_i \times v_j \rightarrow max(N_1, N_2) \rangle$.
However, the meet $(\meet_{O})$ of two octagons, $\langle (v_i \times v_j \rightarrow N_1),
(v_i \times v_j \rightarrow N_2) \rangle$, is always an octagon and is 
computed by taking piece-wise minimum of bounds of corresponding 
octagonal inequalities, $\langle (v_i \times v_j \rightarrow min(N_1, N_2)$.  A closed 
octagon is the smallest octagon following the partial order $\sle_{O}$, among all 
the octagons that abstract the same concrete values.

The octagon domain is a relational abstract domain that permits $2n^2$ 
linear inequalities between $n$ program variables.  The octagonal 
inequalities are of two types: binary or unary inequalities as shown below.
\[Binary: \pm v_i \pm v_j \leq a, v_i \neq v_j \qquad Unary: v_i \leq b, \{a, b\} \in \mathbb{R} \cup \infty \]  
An element in octagon domain, $OctElm$, is a conjunction of such 
inequalities and is called an octagon.  Figure~\ref{octagon} shows an 
example octagon and its associated inequalities.

For building program analyzers using octagon domains, the domain also 
provides few operators like {\em widening ($\nabla$)} and {\em closure ($*$)}.  
The widening operator is used to accelerate convergence for loops in the program
and has a quadratic complexity in the number of variables.  If the bound of an 
octagonal inequality increases every iteration, then $\nabla$ sets the bounds 
to $\infty$.  However, the closure operator is often used to reduce the degree 
of over-approximation resulting from the join operation. Closure always produces 
unique octagonal inequalities, but it is very expensive to compute because of 
cubic complexity in the number of program variables.

The abstract transformer, $apost$, captures the effect of different program 
statements in the abstract domain. The transformer is precise for octagonal 
assignments $(x:=y+1)$ but imprecise for non-octagonal assignments $(x:=y+z)$, 
as shown below.
\[apost_{x:=y+1}(a) = b = \langle x-y \leq 1, y-x \leq 1 \rangle \]  
\[apost_{x:=y+z}(b) = \langle \top \rangle \]

\subsection{Properties of Domain Elements for Decision and Learning}
One of the requirements for a clause learning SAT solver 
is that each element of the partial assignments domain 
needs to have a decomposition into precisely complementable 
elements.  This property of the underlying domain helps to 
guide the search away from the conflicting region of the 
search space.

Thus, clause learning SAT solvers requires a weak complementation property
of the abstraction.  However, numerical abstract domain,  such as 
intervals and octagons are intersections of complementable half-spaces. 
The commonly used library for numerical abstract domains  
is the APRON C library~\cite{apron}.  This library is 
used for the static analysis of the numerical variables 
of a program by abstract interpretation. APRON provides a 
C API interface to various abstract domains and libraries 
such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
{\em Linear Equalities} library.  The aim of such analysis is 
to compute invariants over numerical variables in the 
program~\cite{se2011}.

APRON library is specialized for the abstract interpretation 
framework which does not require complementation 
operator.  To this end, we implement our own template-based 
octagon domain which supports complementation operator.
